---
title: "DataKind 2016 Data Dive: Connecting Public Data Systems to Better Understand System-Involved Youth (Annie E. Casey Foundation)"
author: V. Keith Hughitt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  md_document:
    toc: true
    toc_depth: 2
    variant: markdown_github
---

Introduction
============

Overview
--------

From the project brief, the main goal for the project is to:

>  Demonstrate the benefits of using an Integrated Data System for Child,
>  Youth, and Family (CYF) services by describing the populations of children
>  and youth who have received services from the Behavioral Health and
>  Homelessness programs. Produce population profiles of users of each service
>  and users of both services and make relevant comparisons to the general
>  population of children and youth. Describe the groups of individuals who
>  reappear between and within the same services.

Data
----

Datasets currently being used in this analysis:

- Behavioral Health Services.csv
- CYF Active 2010 to 2016-11-09.csv
- HomelessShelters.csv
- rp_placements_clean.csv

Other datasets and tables available:

- Allegheny County Aggregate Data.xlsx
- IDS Variables.xlsx
- PGHSNAP - Neighborhoods_ All Raw Data.xlsx

Note: some of the files ending with originally ending with "(1).csv" have been
renamed to ".csv".

Data Preparation
================

Load data
---------

```{r load_libraries}
library('readr')
library('dplyr')
library('ggplot2')
library('gplots')
library('knitr')
library('venneuler')
library('viridis')
library('GGally')
library('scales')

set.seed(1)

opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)
options(knitr.duplicate.label='allow')

# version
analysis_version <- 'v1.2'
```

```{r}
# Homeless shelters
shelters <- read_csv('data/raw/HomelessShelters.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_ID_OR_HMIS_CLIENT_ID)

# Remove shelter entries with MCI identifier < 10 digits
# (See note in original xls file)
shelters <- shelters[nchar(shelters$mci) == 10,]

# Behavioral Health Services
bhs <- read_csv('data/raw/Behavioral Health Services.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_UNIQ_ID)

# Child, Youth, and Family (CYF)
cyf <- read_csv('data/raw/CYF Active 2010 to 2016-11-09.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_ID)

# Foster care placements
placements <- read_csv('data/raw/rp_placements_clean.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_ID)
```

Summary of data overlap
-----------------------

```{r dataset_overlap, results='asis'}
# Individuals in all three datasets
common_ids <- intersect(intersect(bhs$mci, cyf$mci), shelters$mci)

# Individuals in both BHS and CYF datasets
bhs_cyf  <- intersect(bhs$mci, cyf$mci)

# Individuals in both CYF and foster care placements
cyf_placements <- intersect(cyf$mci, placements$mci)

# Individuals in both BHS and Shelter datasets
bhs_shelters  <- intersect(bhs$mci, shelters$mci)

# Individuals in both CYF and Shelter datasets
cyf_shelters  <- intersect(cyf$mci, shelters$mci)

# Venn Diagram
bhs_ids <- unique(bhs$mci)
cyf_ids <- unique(cyf$mci)
shelt_ids <- unique(shelters$mci)

vd <- venneuler(c(bhs=length(bhs_ids), cyf=length(cyf_ids),
                  shelt=length(shelt_ids),
                  "bhs&cyf"=length(bhs_cyf),
                  "bhs&shelt"=length(bhs_shelters),
                  "cyf&shelt"=length(cyf_shelters),
                  "bhs&cyf&shelt"=length(common_ids)))
vd$labels <- c(paste0("bhs\n", length(bhs_ids)),
               paste0("cyf\n", length(cyf_ids), "\n(", length(cyf_placements), 
                      " foster placements)"),
               paste0("shelt\n", length(shelt_ids)))
plot(vd)

# Overlap statistics
cat(sprintf('- Overlap between BHS & CYF: %d\n', length(bhs_cyf)))
cat(sprintf('- Overlap between BHS & Shelter: %d\n', length(bhs_shelters)))
cat(sprintf('- Overlap between CYF & Shelter: %d\n', length(cyf_shelters)))
cat(sprintf('- Overlap between all three: %d\n', length(common_ids)))
```

Data overview
-------------

### Behavior and Health Services

```{r, bhs_hist}
dim(bhs)
range(table(bhs$mci))
hist(table(bhs$mci), main='Number of BHS entries per individual')
```

### CYF

```{r, cyf_hist}
dim(cyf)
range(table(cyf$mci))
hist(table(cyf$mci), main='Number of homeless shelter entries per individual')
```

### Homeless shelters

```{r, shelters_hist}
dim(shelters)
range(table(shelters$mci))
hist(table(shelters$mci), main='Number of homeless shelter entries per individual')
```

Preview of data for a single individual
---------------------------------------

```{r results='asis'}
# counts per dataset (shared across all three)
data_counts <- merge(merge(
    bhs %>% group_by(mci) %>% summarize(num_bhs=n()),
    cyf %>% group_by(mci) %>% summarize(num_cyf=n()), by='mci'),
    shelters %>% group_by(mci) %>% summarize(num_shelter=n()), by='mci')

# individuals with multiple rows in all three tables
multi_row_individuals <- data_counts[apply(data_counts, 1, min) > 1,]$mci

# random individual with multiple entries in each table
example_youth <- multi_row_individuals[1]
```

### BHS

```
bhs %>% 
    filter(mci == example_youth) %>% 
    head(10) %>%
    kable
```

### CYF

```{r results='asis'}
cyf %>% 
    filter(mci == example_youth) %>% 
    kable
```

### Shelters

```{r results='asis'}
shelters %>% 
    filter(mci == example_youth) %>% 
    kable
```

Results
=======

Below, we will explore some of the properties of intersections of two or three
of the datasets.

BHS, CYF, and Shelters
----------------------

```{r}
ids_to_use <- common_ids
analysis_name <- 'combined'

datasets <- c('bhs', 'cyf', 'shelters')
```

```{r child='src/results.Rmd'}
```

### Race vs. other variables

```{r}
# is there a relationship between race and variables? (just considered dominant
# categories)
dat_subset <- dat %>% 
    filter(RACE %in% c('White', 'Black or African American')) %>%
    select(-mci)
ggpairs(dat_subset, mapping=aes(colour=RACE, alpha=0.4))
```

### Common behavior and health service types

#### By Gender

```{r}
# top 5 bhs types
top_bhs_types <- names(sort(table(bhs$SERVICE_NAME), decreasing=TRUE)[1:5])

# create a modified version of the combined dataset with bhs type columns
dat_with_bhs_type <- merge(dat, bhs %>% select(mci, SERVICE_NAME), by='mci')

dat_with_bhs_type$SERVICE_NAME[!dat_with_bhs_type$SERVICE_NAME %in% top_bhs_types] <- 'Other'

# vs. gender
ggplot(dat_with_bhs_type, aes(x=GENDER)) +
    geom_bar(aes(fill=SERVICE_NAME), position="dodge")
```

#### By Race

```{r bhs_vs_race}
#x <- dat_with_bhs_type %>% 
#    group_by(GENDER, SERVICE_NAME) %>% 
#    summarize(n=n()) %>%
#    mutate(freq = n / sum(n))

# vs. race
ggplot(dat_with_bhs_type %>% filter(RACE %in% c('White', 'Black or African American')), aes(x=RACE)) +
    geom_bar(aes(fill=SERVICE_NAME))
```

Shelters, CYF
-------------

```{r}
ids_to_use <- cyf_shelters
analysis_name <- 'cyf_shelters'

datasets <- c('shelters', 'cyf')
```

```{r child='src/results.Rmd'}
```

BHS, CYF
--------

```{r}
ids_to_use <- bhs_cyf
analysis_name <- 'bhs_cyf'

datasets <- c('bhs', 'cyf')
```

```{r child='src/results.Rmd'}
```

System Info
===========

```{r}
sessionInfo()
```
