---
title: "DataKind 2016: Annie E. Casey Foundation"
author: V. Keith Hughitt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  md_document:
    variant: markdown_github
---

Overview
========

...

Data
----

### Allegheny County Aggregate Data.xlsx

### Behavioral Health Services.csv

### CYF Active 2010 to 2016-11-09.csv

### HomelessShelters(1).csv

### IDS Variables.xlsx

### PGHSNAP - Neighborhoods_ All Raw Data.xlsx

### rp_placements_clean.csv

Note: some of the files ending with "(1).csv" have been renamed to ".csv".

Data Preparation
================

```{r load_libraries}
library('readr')
library('dplyr')
library('ggplot2')
library('gplots')
library('knitr')
library('venneuler')
library('viridis')

opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)
```

```{r}
# Homeless shelters
shelters <- read_csv('data/raw/HomelessShelters.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_ID_OR_HMIS_CLIENT_ID)

# Remove shelter entries with MCI identifier < 10 digits
# (See note in original xls file)
shelters <- shelters[nchar(shelters$mci) == 10,]

# Behavioral Health Services
bhs <- read_csv('data/raw/Behavioral Health Services.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_UNIQ_ID)

# CYF Active
cyf <- read_csv('data/raw/CYF Active 2010 to 2016-11-09.csv.gz', progress=FALSE) %>%
    rename(mci=MCI_ID)
```

```{r results='asis'}
# Individuals in all three datasets
common_ids <- intersect(intersect(bhs$mci, cyf$mci), shelters$mci)

# Venn Diagram
bhs_ids <- unique(bhs$mci)
cyf_ids <- unique(cyf$mci)
shelt_ids <- unique(shelters$mci)

vd <- venneuler(c(bhs=length(bhs_ids), cyf=length(cyf_ids),
                  shelt=length(shelt_ids),
                  "bhs&cyf"=length(intersect(bhs_ids, cyf_ids)),
                  "bhs&shelt"=length(intersect(bhs_ids, shelt_ids)),
                  "cyf&shelt"=length(intersect(cyf_ids, shelt_ids)),
                  "bhs&cyf&shelt"=length(intersect(bhs_ids, intersect(cyf_ids, shelt_ids)))))
vd$labels <- c(paste0("bhs\n", length(bhs_ids)),
               paste0("cyf\n", length(cyf_ids)),
               paste0("shelt\n", length(shelt_ids)))
plot(vd)

# Overlap statistics
cat(sprintf('- Overlap between BHS & CYF: %d\n', length(intersect(bhs_ids, cyf_ids))))
cat(sprintf('- Overlap between BHS & Shelter: %d\n', length(intersect(bhs_ids, shelt_ids))))
cat(sprintf('- Overlap between CYF & Shelter: %d\n', length(intersect(cyf_ids, shelt_ids))))
cat(sprintf('- Overlap between all three: %d\n', length(intersect(bhs_ids, intersect(cyf_ids, shelt_ids)))))

# Drop rows for individuals with missing data
cat(sprintf('- Dropping %d / %d rows from shelter dataset (not shared)\n', 
            sum(!shelters$mci %in% common_ids),
            nrow(shelters)))
shelters <- shelters[shelters$mci %in% common_ids,]

cat(sprintf('- Dropping %d / %d rows from behavior dataset (not shared)\n', 
            sum(!bhs$mci %in% common_ids),
            nrow(bhs)))
bhs <- bhs[bhs$mci %in% common_ids,]

cat(sprintf('- Dropping %d / %d rows from CYF dataset (not shared)\n', 
            sum(!cyf$mci %in% common_ids),
            nrow(cyf)))
cyf <- cyf[cyf$mci %in% common_ids,]
# Placements
#placements <- read_csv('data/raw/rp_placements_clean.csv', row_names=FALSE)

# counts per dataset
data_counts <- merge(merge(
    bhs %>% group_by(mci) %>% summarize(num_bhs=n()),
    cyf %>% group_by(mci) %>% summarize(num_cyf=n()), by='mci'),
    shelters %>% group_by(mci) %>% summarize(num_shelter=n()), by='mci')

# individuals with multiple rows in all three tables
multi_row_individuals <- data_counts[apply(data_counts, 1, min) > 1,]$mci
```

Exploratory Data Analaysis
==========================

### Behavior and Health Services

#### Overview

```{r, bhs_hist}
dim(bhs)
range(table(bhs$mci))
hist(table(bhs$mci), main='Number of BHS entries per individual')
```

```{r, results='asis'}
kable(head(bhs))
```

#### Single Individual

```{r results='asis'}
# random individual with multiple entries in each table
example_youth <- multi_row_individuals[1]

bhs %>% 
    filter(mci == example_youth) %>% 
    head(10) %>%
    kable
```

### Homeless shelters

```{r, shelters_hist}
dim(shelters)
range(table(shelters$mci))
hist(table(shelters$mci), main='Number of homeless shelter entries per individual')
```

```{r, results='asis'}
kable(head(shelters))
```

#### Single Individual

```{r results='asis'}
shelters %>% 
    filter(mci == example_youth) %>% 
    kable
```

### CYF

```{r, cyf_hist}
dim(cyf)
range(table(cyf$mci))
hist(table(cyf$mci), main='Number of homeless shelter entries per individual')
```

```{r, results='asis'}
kable(head(cyf))
```

#### Single Individual

```{r results='asis'}
cyf %>% 
    filter(mci == example_youth) %>% 
    kable
```

Combine Datasets
================

```{r}
# Behavior and health services
bhs_collapsed <- bhs %>%
    group_by(mci) %>%
    summarize(num_visits=n(),
              num_unique_providers=n_distinct(PRVDR_NAME),
              #county_tot=sum(CNTY_TOT),
              total_units=sum(TOT_UNITS))

# drop INVLV and birthdate columns
cyf_collapsed <- cyf %>% 
    select(-CAS_ID, -CL_ID, -BRTH_DT, -starts_with('INVLV'))

# Just keep one row for each individual
cyf_collapsed <- cyf_collapsed[!duplicated(cyf_collapsed$mci),]

# Shelters
shelters_collapsed <- shelters %>%
    group_by(mci) %>%
    summarize(num_shelter_stays=n(),
              num_unique_hud_proj_types=n_distinct(HUD_PROJECT_TYPE),
              min_stay=min(DURATION_OF_STAY),
              avg_stay=median(DURATION_OF_STAY),
              max_stay=max(DURATION_OF_STAY))

# Combined dataset
dat <- merge(merge(bhs_collapsed, cyf_collapsed, by='mci'), 
             shelters_collapsed, by='mci')

# Drop the two rows with NA's
dat <- dat[complete.cases(dat),]
```

Variable correlation heatmap

```{r, variable_heatmap}
# version of data with categorical variabled numerically encoded
dat_numeric <- dat
dat_numeric$RACE <- as.numeric(factor(dat$RACE))
dat_numeric$GENDER <- as.numeric(factor(dat$GENDER))

# drop id column
dat_numeric <- dat_numeric[,2:ncol(dat_numeric)]
dat_numeric <- as.matrix(dat_numeric)

cor_mat <- cor(dat_numeric, method='spearman')
heatmap.2(cor_mat, trace='none', col=viridis)
```

Individual correlation heatmap

```{r, individual_heatmap}
cor_mat <- cor(t(dat_numeric), method='spearman')
heatmap.2(cor_mat, trace='none', col=viridis)
```

System Info
===========

```{r}
sessionInfo()
```
